\documentclass[11pt,fleqn]{article}
\usepackage{amsmath}
\usepackage{changepage}% http://ctan.org/pkg/changepage
\numberwithin{equation}{subsection}
\usepackage[parfill]{parskip}
\setlength{\marginparwidth}{0pt}
\usepackage{layout}
\begin{document}
\title{Aetherling Nodes and Example Pipelines}

\section{TODOS}
\begin{enumerate}
    \item Need to add area/time estimates for each element and application
\end{enumerate}

\section{Type Notation}

\subsection{Combinational Element Types}
The type signature of combinational elements is:
\begin{adjustwidth}{2.5 em}{0 pt}
    input 0 type $\rightarrow$ ... $\rightarrow$ input n-1 type $\rightarrow$ output type
\end{adjustwidth}
Since these combinational elements process all inputs and output every 
clock cycle, they do not need ready-valid signals to indicate when they are
ready for input or emitting valid outputs.

Each input or output has type T or T[p]. T is a base type, and it can contain 
nested types like arrays that are not relevant for the current operation. T[p] 
is an array of length p of T's. T[p][q] is an array of length p of arrays of 
length q of T's.

\subsection{Sequential Element Types}
The type signature of sequential elements is:
\begin{adjustwidth}{2.5 em}{0 pt}
    \{input 0 num cycles, input 0 type per cycle\} $\rightarrow$ ...
    $\rightarrow$ \\
    \{input n-1 num cycles, input n-1 type per cycle\} $\rightarrow$ \\ 
    \{ouput num cycles, output type per cycle\}
\end{adjustwidth}

One of the inputs or outputs of a stream may have different types during different
clock cycles in a stream. For example, a sequential reduce may emit invalid 
data for most of the stream's cycles and then emit the result on the final
cycle of the stream. Instead of \{cycles, type for all cycles\}, 
this type is represented in the following way:
\begin{adjustwidth}{2.5 em}{0 pt}
    [type at cycle 0, type at cycle 1, ..., type at cycle n-1]
\end{adjustwidth}
For short-hand where a type is the same for multiple cycles:
\begin{adjustwidth}{2.5 em}{0 pt}
    [type at cycle 0(0:n-2), type at cycle n-1]
\end{adjustwidth}
The invalid type is represented as $\emptyset$.

These elements interfaces also must have clock inputs and may have ready-valid 
inputs and outputs. These ready-valid handshake ports indicate:
\begin{enumerate}
        \item ready
            \subitem input: indicates to this sequential element that the next 
            one in the pipeline has completed its prior input stream and is 
            ready to receive more input.
            \subitem output: indicates to the previous sequential element in the 
            pipeline that this one has completed its prior input stream and is
            ready to receive more input.
        \item valid
            \subitem input: indicates to this sequential element that the previous
            one in the pipeline is emitting valid data that this sequential 
            element can use as input.
            \subitem output: indicates to the next sequential element in the
            pipeline that this one is emitting valid data that the next one
            can use as input.
\end{enumerate}

\section{Basic Elements}
These are elements that are not built using other Aetherling elements.

\subsection{Combinational Elements}

\begin{enumerate}
    \item tuple :: S $\rightarrow$ T $\rightarrow$ (S,T)
    \item lb p w :: T[p] $\rightarrow$ T[w+p-1]
    \item overlap\_partition p w :: T[w+p-1] $\rightarrow$ T[p][w]
    \item partition p k :: T[k] $\rightarrow$ T[k/p][p]
    \item flatten p k :: T[k/p][p] $\rightarrow$ T[k]
    \item map p f :: $S_1$[p] $\rightarrow$ ... $\rightarrow$ $S_{n-1}$[p] $\rightarrow$ 
        T[p]
        \subitem s.t. f :: $S_1 \rightarrow$ ... $\rightarrow S_n \rightarrow$ T 
    \item reduce p f :: T[p] $\rightarrow$ T
        \subitem s.t. f :: (T,T)$\rightarrow$ T
    \item up k :: T $\rightarrow$ T[k]
    \item down k :: T[k] $\rightarrow$ T
    \item zip :: (S[k],T[k]) $\rightarrow$ (S,T)[k]
    \item unzip :: (S,T)[k] $\rightarrow$ (S[k], T[k])
    \item mem\_read p :: () $\rightarrow$ T[p]
    \item mem\_write p :: T[p] $\rightarrow$ ()
\end{enumerate}

\subsection{Sequential Elements}

\subsubsection{Sequential Versions Of Basic Combinational Elements}
These are the sequential elements that cannot be implemented using the
combinational elements. In an ideal world, there would only be basic
combinational logic blocks and rate changing sequential elements, but these
elements are better implemented using logical blocks that do not belong in
Aetherling.
\begin{enumerate}
    \item reduce\_seq p f :: [T[p], $\emptyset$(1:p-1)] $\rightarrow$ [$\emptyset$(0:p-2), T]
        \subitem s.t. f :: (T,T)$\rightarrow$ T
    \item reduce\_seq\_stream p f :: \{p, T\} $\rightarrow$ [$\emptyset$(0:p-2), T]
        \subitem s.t. f :: (T,T)$\rightarrow$ T
    \item up\_seq k :: [T, $\emptyset$(1:k-1)] $\rightarrow$ \{k, T\}
    \item down\_seq k :: {k, T} $\rightarrow$ [T, $\emptyset$(1:k-1)]
\end{enumerate}

\subsubsection{Rate Changing Elements}

\begin{enumerate}
    \item serialize k :: [T[k], $\emptyset$(1:k-1)] $\rightarrow$ \{k, T\}
    \item deserialize k :: \{k, T\} $\rightarrow$ [$\emptyset$(0:k-2), T[k]]
\end{enumerate}

\section{Basic Applications}
These are simple combinations of the basic elements.

\subsection{Passthrough}
\begin{enumerate}
    \item mem\_write 1 \$ mem\_read 1
    \item mem\_write t \$ mem\_write t 
\end{enumerate}

\subsection{Array-Stream Conversions}
\begin{enumerate}
    \item mem\_write 1 \$ deserialize t \$ mem\_read t
        \subitem Note that mem\_read fires onces every t'th clock cycle
    \item mem\_write 1 \$ deserialize t \$ serialize t \$ mem\_read 1
        \subitem Note mem\_write fires every t'th clock cycle
\end{enumerate}

\subsection{Map}
\begin{enumerate}
    \item mem\_write 1 \$ map 1 (+1) \$ mem\_read 1
    \item mem\_write t \$ map t (+1) \$ mem\_read t
    \item mem\_write t \$ map t (+1) \$ map t f1 \$ mem\_read t
\end{enumerate}

\subsection{Reduce}
\begin{enumerate}
    \item mem\_write 1 \$ reduce t (+) \$ mem\_read t
    \item mem\_write 1 \$ reduce t (+) \$ deserialize t f \$ mem\_read 1 
        \subitem Note everything after deserialize fires every t'th clock cycle
    \item mem\_write 1 \$ reduce\_seq t (+) \$ mem\_write t
        \subitem Note mem\_write fires every t'th clock cycle
    \item mem\_write 1 \$ reduce\_seq\_stream t (+) \$ mem\_write 1
\end{enumerate}

\subsection{Array Dimension Conversions}
\begin{enumerate}
    \item mem\_write $\frac{t}{p}$ \$ partition p t \$ mem\_read t
        \subitem Note that the element type mem\_write is writing is T[p], 
        and it writes $\frac{t}{p}$ of them every clock.
    \item mem\_write t \$ flatten t \$ partition t \$ mem\_read t 
    \item mem\_write 1 \$ flatten t \$ partition t \$ deserialize t \$ mem\_read t
        \subitem Note everything after deserialize fires onces every t'th 
        clock cycle
    \item mem\_write 1 \$ down t \$ up t \$ mem\_read 1
    \item mem\_write 
\end{enumerate}

\section{Composed Elements}
These are elements that are composed from basic elements and other other composed
elements.

\subsection{Sequential Versions Of Basic Combinational Elements}
These are operations that perform the same operations as the basic ones, but
do it over multiple clock cycles.
\begin{enumerate}
    \item map\_seq p f :: [$S_1$[p], $\emptyset$(1:p-1)] $\rightarrow$ ...
        $\rightarrow$ [$S_{n-1}$[p], $\emptyset$(1:p-1)] $\rightarrow$
        [$\emptyset$(0:p-2), T[p]]
        \subitem s.t. f :: $S_1 \rightarrow$ ... $\rightarrow S_n \rightarrow$ T
        \subitem This map takes all the inputs in on the first cycle of the 
        stream and emits all the outputs on the final cycle of the stream.
        \subitem implementation: map\_seq p f = deserialize p \$ map 1 f \$ \\ serialize p
        \subitem note that in the above implementation, the type for serialize 
        contains all the different input types to map
    \item map\_seq\_stream p f :: [$S_1$[p], $\emptyset$(1:p-1)] $\rightarrow$ ...
        $\rightarrow$ [$S_{n-1}$[p], $\emptyset$(1:p-1)] $\rightarrow$
        \{p, T\}
        \subitem s.t. f :: $S_1 \rightarrow$ ... $\rightarrow S_n \rightarrow$ T
        \subitem This map takes all the inputs in on the first cycle of the 
        stream and emits one output on each cycle of the stream.
        \subitem implementation: map\_seq\_stream p f = map 1 f \$ serialize p
    \item map\_part\_par p k f :: []
\end{enumerate}

\end{document}
